Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-03-17 paper/manuscript.Rmd
Update manuscript.Rmd

For several decades after their introduction to Fogo Island, the caribou
population did not exceed ~100 individuals [@Bergerud1989, Newfoundland and
Labrador Wildlife Division, unpublished data]. During the 1990s, population
density reached 300 animals and it is unknown when caribou began to occupy other
islands, anecdotal evidence suggests it was during the 1990s. We surmise that as
competition among conspecifics increased along with population size,
density-dependent habitat selection resulted in expansion of the population to
nearby islands. Following Ideal Free Distribution theory, caribou should swim to
new islands when the average fitness of individuals on the starting island
exceeds the density-fitness equilibrium [@Morris1987]. While this is an ultimate
explanation to a series of proximate observations, fine-scale density-dependent
habitat selection is theoretically possible [@Webber2018]. Given the presumed
low density of coyotes on Fogo Island and the minimal risk they pose, predation risk
does not appear to be a major driver of swimming.

>>>>

For several decades after their introduction to Fogo Island, the caribou
population did not exceed ~100 individuals [@Bergerud1989, Newfoundland and
Labrador Wildlife Division, unpublished data]. During the 1990s, population
density reached 300 animals and it is unknown when caribou began to occupy other
islands, anecdotal evidence suggests it was during the 1990s. We surmise that as
competition among conspecifics increased along with population size,
density-dependent habitat selection resulted in expansion of the population to
nearby islands. Following Ideal Free Distribution theory, caribou should swim to
new islands when the average fitness of individuals on the starting island
exceeds the density-fitness equilibrium [@Morris1987]. While this is an ultimate
explanation, our data suggest caribou only remain on smaller islands for a few
days at a time, so the fitness equilibrium remains a theoretical construct. More
likely, however, is that fine-scale competition, density-dependent habitat
selection, and forage depletion drives individuals to peridiodically swim
between islands. Given the presumed low density of coyotes on Fogo Island and
the minimal risk they pose, predation risk does not appear to be a major driver
of swimming.
